Search-based Synthesis of Network Intrusion Detection Rules

### 2019-09-15 paper/paper.tex
pass

\usepackage{booktabs}
\usepackage{cite}
\usepackage{url}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
\usepackage{listings}
\usepackage[caption=false, font=footnotesize]{subfig}
%%%%%%%%%%%%% code listing
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\scriptsize\ttfamily\bfseries,
  language=C,             % choose the language of the code
  frame=single,              % adds a frame around the code
  aboveskip=0pt,
  belowskip=0pt,
  breaklines=true,           % sets automatic line breaking
  breakatwhitespace=false,   % sets if automatic breaks should only happen at
  showspaces=false,
  %numbersep=5pt,              % Abstand der Nummern zum Text
  %tabsize=2,                  % Groesse von Tabs
  %extendedchars=true,         %
  %breaklines=true,            % Zeilen werden Umgebrochen
  keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
}
\usepackage{balance}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

@@

Search-based algorithms need to characterize the problem (\ie{}, a
solution candidate) in some way. In this case, the problem is the
rule.  \tname\ encodes the rule options segment (see
Section~\ref{sec:example-suricata-rules}) of a Suricata rule as a
vector of Option objects. For example, \tname{} models the option
\CodeIn{flags} from the example on Figure~\ref{fig:synflood-example}
with a data structure storing two fields, one of character type and
another of integer type. This enables \tname{} to mutate each element
separately.\Mar{Lucas, por favor, confirme.}

>>>>

\usepackage{booktabs}
\usepackage{cite}
\usepackage{url}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{matrix,fit,shapes,calc,positioning,shadows,arrows,shapes,backgrounds,decorations.markings,fadings}
\usepackage{listings}
\usepackage[caption=false, font=footnotesize]{subfig}
%%%%%%%%%%%%% code listing
\renewcommand{\ttdefault}{pcr}
\lstset{
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\scriptsize\ttfamily\bfseries,
  language=C,             % choose the language of the code
  frame=single,              % adds a frame around the code
  aboveskip=0pt,
  belowskip=0pt,
  breaklines=true,           % sets automatic line breaking
  breakatwhitespace=false,   % sets if automatic breaks should only happen at
  showspaces=false,
  %numbersep=5pt,              % Abstand der Nummern zum Text
  %tabsize=2,                  % Groesse von Tabs
  %extendedchars=true,         %
  %breaklines=true,            % Zeilen werden Umgebrochen
  keywords=[2]{tcp, flag, threshold, track, count, seconds, classtype, sid}
}
\usepackage{balance}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}

@@

One important step in search-based optimization is to characterize a
candidate solution to the target problem. Candidate solutions are the
objects that the search tries to optimize. In this case, a rule is a
candidate solution.  \tname\ encodes the rule options segment of a
Suricata rule (see Section~\ref{sec:example-suricata-rules}) as a
vector of Option objects. Some options are discarded as they only
serve for documentation. For example, the options \CodeIn{msg},
\CodeIn{classtype}, and \CodeIn{miscactivity} have no influence in
triggering the action of rules. The option \CodeIn{flags}, appearing
in the example on Figure~\ref{fig:synflood-example}, is relevant to
trigger the action. \tname{} models that option with a data structure
storing two fields, one of character type and another of integer
type. This enables \tname{} to mutate each element
separately.

Options can be specific to a given protocol or general and they can be
influential on triggering an action or
not. Figure~\ref{fig:distribution-rules-protocol} shows the
distributions of these options per protocol and their relevance for
triggering rule actions.

\pgfplotsset{width=6cm,compat=1.8}
\begin{figure}[h!]
  \centering
  \begin{tikzpicture}
    \begin{axis}[
        ybar stacked,
        enlargelimits=0.15,
        legend style={at={(0.5,-0.15)},
          anchor=north,legend columns=-1},
        ylabel={\#rules},
        symbolic x coords={gen, http, icmp, udp, tcp},
        xtick=data,
        nodes near coords,
        nodes near coords align={vertical},
      ]
      \addplot coordinates {(gen,70) (http,34) (icmp,94) (udp,84) (tcp,43)};
      \addplot coordinates {(gen,50) (http,20) (icmp,134) (udp,90) (tcp,123)};    
      \legend{relevant, irrelevant}
    \end{axis}
  \end{tikzpicture}
  \caption{\label{fig:distribution-rules-protocol}Distribution of
    rules per protocol and relevance for triggering action.}
\end{figure}

To bootstrap the search, \tname{} needs to select an initial
solution. When creating such candidate solution, \tname{} first looks
for the protocol related to the malicious input traffic. Then, it
selects the relevant options for that protocol and initializes the
option vector with default values. \Mar{Suricata define valores
  default?  Vcs. definiram?  Isto foi baseado em popularidade ou o
  que?}
